### 2011-04-24 doc/Manual/BaconManual.tex
Started on paper.

\documentclass{article}
\begin{document}

\section{Introduction}

Bacon is a front end for OpenCL that provides extra convienience compared
to the raw OpenCL API and OpenCL C programming language.

Bacon is a data parallel extension to OpenCL C that adds the following
features:

 - Native 2d and 3d arrays with multi-dimensional addressing syntax.

 - Images as multi-dimensional arrays.

 - Support for "printf debugging".

 - Automatic C++ wrapper generation.

\section{Multi-Dimentional Arrays}

Support is provided for native 2D and 3D arrays. Some examples:

Declare a 2D array with eight rows and four columns:
$$\hbox{\tt array2d<int> table[8, 4];}$$

Assign the value 8 to the third row, second column of the array.
$$\hbox{\tt table[2, 1] = table.rows;}$$

\end{document}

### 2011-04-24 doc/Paper/Paper.tex
Started on paper.

\documentclass{llncs}

\begin{document}

\title{Bacon: GPU Programming With Just in Time Specialization}

\author{Nat Tuck\inst{1}}

\institute{University of Massachusetts Lowell, Lowell MA 01854, USA}

\maketitle

\begin{abstract}

This paper describes Bacon, a programming language that provides
simpler development and better optimizations for OpenCL compatible
general purpose GPUs.

\end{abstract}

\section{Introduction}

OpenCL is a standard language for data parallel computation primarily
targetted at graphics processing units.

Although OpenCL maps well to GPU hardware and provides detailed
control, it has some weaknesses. The C++ interface to OpenCL is very
verbose, and requires quite a bit of boilerplate to get any program to
run at all. OpenCL doesn't provide any mechanism for dynamic memory
allocation. Existing implementations of OpenCL don't do loop unrolling
for loops with unknown iteration counts.

Bacon was developed to try to improve the OpenCL programming
experience. It extends the OpenCL C programming language and provides
a compiler that hopes to solve these issues. C++ wrapper code is
generated automatically. Just in time specialization allows the 
illusion of dynamic stack allocation as well as the unrolling of
some loops.

Partial evaluation has been shown to provide improved performance in
scientific computing applications. A good example of this is shown in
\cite{Berlin:1990}.

The design of the OpenCL runtime system depends on device-specific
code generation occuring at runtime, which suggests the use of
parameter information for optimization at runtime.

\section{Bacon}
\subsection{Language}

The Bacon programming language is based on OpenCL C with a small
number of modifications.

-

\subsection{Implementation}

Hai

\section{Result}

Hai

\section{Conclusion}

Hai

\bibliography{bacon}{}
\bibliographystyle{splncs03}
\end{document}

### 2011-04-29 doc/Paper/Paper.tex
Working on the paper.

@@
 \documentclass{llncs}
+\usepackage{graphicx}
+\usepackage{float}
 
@@
 
-\title{Bacon: GPU Programming With Just in Time Specialization}
+\title{Bacon: A GPU Programming Language With Just in Time Specialization}
 
-\author{Nat Tuck\inst{1}}
+\author{Nat Tuck}
 
@@
 
-OpenCL is a standard language for data parallel computation primarily
-targetted at graphics processing units.
+The use of Graphics Processing Units (GPUs) for general purpose
+parallel computing has become increasingly feasible over the last few
+years. In response to the platform specific programming solutions from
+Nvidia and Microsoft, Apple developed the OpenCL standard as an open and
+cross platform programming interface to this hardware.
 
-Although OpenCL maps well to GPU hardware and provides detailed
-control, it has some weaknesses. The C++ interface to OpenCL is very
-verbose, and requires quite a bit of boilerplate to get any program to
-run at all. OpenCL doesn't provide any mechanism for dynamic memory
-allocation. Existing implementations of OpenCL don't do loop unrolling
-for loops with unknown iteration counts.
+The OpenCL standard consists of two major pieces. First, it defines a
+programming language called OpenCL C for writing compute kernels to
+run on parallel hardware. Second, it defines runtime APIs for C and
+C++ that allow these kernels to be compiled, loaded, and executed from
+programs running on a host machine.
+
+OpenCL C uses the syntax of C99 and provides a set of built in data
+types and functions that expose the numeric computatation capabilities
+common to modern GPU devices. The specification explicitly disallows
+the use of various C99 functionality that is not supported by GPU
+hardware, including function pointers, recursion, and any sort of
+dynamic memory allocation or array sizing.
+
+Rather than providing a stand-alone program to compile OpenCL C
+kernels, the OpenCL C and C++ APIs give developers the nessisary
+pieces to build a compiler into their host program. This allows OpenCL
+programs to be portable across different hardware archetectures by
+delaying compilation until runtime when the target GPU device is
+known, but requires each developer to write quite a bit of code to
+load the source code and perform various other bookkeeping.
 
@@
 Partial evaluation has been shown to provide improved performance in
-scientific computing applications. A good example of this is shown in
+scientific computing applications. A good example of this is shown by Berlin
 \cite{Berlin:1990}.
@@
 The Bacon programming language is based on OpenCL C with a small
-number of modifications.
+number of modifications:
 
--
+- Kernel definition format that allows auto-generation of
+  C++ interface code.
+- Native 2D and 3D array syntax.
+- Runtime value specialization, allowing for
+  - Variable size array declarations
+  - The expectation of loop unrolling
+
+Because private array items are actually assigned to device registers,
+at least in the AMD implementation of OpenCL, this means that a loop
+over a variable sized local array can be expected to be as efficient
+as the manually unrolled version.
 
@@
 
-Hai
+Bacon consists of a compiler that translates Bacon C into an
+intermediate representation and generates interface code to load and
+execute that code at runtime from a C++ host application.
 
-\section{Result}
+\section{Performance Results}
 
-Hai
+\subsection{Matrix Multiplication}
+
+To analyze the performance of the Bacon system, we performed a number
+of tests multiplying random 4096x4096 matrixes of single precision
+point numbers.
+
+Testing was performed using the hardware shown in Table~\ref{hw}. At
+the time of this writing this represents an inexpensive (\~\$500) new
+PC for computer gaming, but both the CPU and GPU archetectures are
+less than one design generation behind the latest hardware from AMD.
+
+\begin{table}[htb]
+\begin{tabular}{ l @{\hspace{10pt}} l @{\hspace{10pt}} l }
+Device & Speed & RAM \\
+\hline
+\noalign{\smallskip}
+AMD Phenom II X3 720 & 3 Cores @ 2.8 GHz & 4GB DDR2 @ 800MHz \\
+\noalign{\smallskip}
+ATI Radeon 5830 & 1120 Stream Cores @ 800 MHz & 1GB GDDR5 @ 1000 MHz \\
+\noalign{\smallskip}
+\end{tabular}
+\caption{Hardware used for performance testing}\label{hw}
+\end{table}
+
+The baseline measurement is a naive matrix multiplication written in OpenCL C.
+
+\begin{table}
+\begin{tabular}{ l @{\hspace{10pt}} r @{\hspace{10pt}} r }
+Test & Time (s) & Speedup \\
+\hline
+\noalign{\smallskip}
+OpenCL - Naive & 11.9 & 1.0 \\
+\noalign{\smallskip}
+OpenCL - Hand Vectorized & 2.54 & 4.7 \\
+\noalign{\smallskip}
+Bacon - Naive (Best) & 3.45 & 3.5 \\
+\noalign{\smallskip}
+Bacon - Blocked (Best) & 1.97 & 6.1 \\
+\noalign{\smallskip}
+\end{tabular}
+\caption{Summary of 4k Matrix Multiplication Performance}\label{mm1}
+\end{table}
+
+\begin{figure}[htb]
+\begin{center}
+\includegraphics[clip]{unrolling}
+\end{center}
+\caption{Summary of 4k Matrix Multiplication Performance}\label{unroll}
+\end{figure}
 

### 2011-04-29 examples/BlockMatMul/bmmul.cc

@@ -43,9 +43,12 @@ run_test(string c_file, string a_file, string b_file, int block_size, bool priv)
 }
 
 void
-random_test(int nn, bool check, int block_size, bool priv)
+random_test(int nn, bool check, int block_size, bool priv, bool print_time)
 {
     BlockMatMul mmul;
+
+    Bacon::Timer tt;
+    double seconds;
 
     Bacon::Array2D<float> aa(nn, nn);
     aa.fill_random();
@@ -58,11 +61,26 @@ random_test(int nn, bool check, int block_size, bool priv)
 
     Bacon::Array2D<float> cc;
 
+    tt.reset();
     if (priv)
         cc = mmul.blocked_mat_mul_private(aa, bb, block_size);
     else
         cc = mmul.blocked_mat_mul_local(aa, bb, block_size);
-   
+    seconds = tt.time();
+
+    if (print_time) {
+        cout << "First run took " << seconds << " seconds." << endl;
+
+        tt.reset();
+        if (priv)
+            cc = mmul.blocked_mat_mul_private(aa, bb, block_size);
+        else
+            cc = mmul.blocked_mat_mul_local(aa, bb, block_size);
+        seconds = tt.time();
+
+        cout << "Second run took " << seconds << " seconds." << endl;
+    }
+
     if(!check) {
         cout << "Result not checked." << endl;
         return;
@@ -100,11 +118,12 @@ main(int argc, char* argv[])
 
     bool check_result = false;
     bool private_mem  = false;
+    bool print_time   = false;
 
     int random_size = 0;
     int block_size = 1;
 
-    while ((opt = getopt(argc, argv, "hpca:b:o:n:k:")) != -1) {
+    while ((opt = getopt(argc, argv, "hpcta:b:o:n:k:")) != -1) {
         switch(opt) {
         case 'a':
             a_file = string(optarg);
@@ -127,6 +146,9 @@ main(int argc, char* argv[])
         case 'p':
             private_mem = true;
             break;
+        case 't':
+            print_time = true;
+            break;
         case 'h':
             show_usage();
             return 0;
@@ -138,7 +160,7 @@ main(int argc, char* argv[])
 
 
     if (random_size != 0) {
-        random_test(random_size, check_result, block_size, private_mem);
+        random_test(random_size, check_result, block_size, private_mem, print_time);
     }
     else {
         run_test(c_file, a_file, b_file, block_size, private_mem);

### 2011-04-29 examples/MatMul/mmul.cc

@@ -10,6 +10,8 @@ using std::string;
 #include <stdlib.h>
 #include <unistd.h>
 
+#include <Bacon/Timer.hh>
+
 #include "gen/MatMul.hh"
 
 void
@@ -38,9 +40,12 @@ run_test(string c_file, string a_file, string b_file)
 }
 
 void
-random_test(int nn, bool check = true)
+random_test(int nn, bool check = true, bool timing = false)
 {
     MatMul mmul;
+
+    Bacon::Timer tt;
+    double seconds;
 
     Bacon::Array2D<float> aa(nn, nn);
     aa.fill_random();
@@ -48,8 +53,21 @@ random_test(int nn, bool check = true)
     Bacon::Array2D<float> bb(nn, nn);
     bb.fill_identity_matrix();
 
+    tt.reset();
     Bacon::Array2D<float> cc = mmul.mat_mul(aa, bb);
-   
+    seconds = tt.time();
+
+    if (timing) {
+        cout << "First run took " << seconds << " seconds." << endl;
+
+        tt.reset();
+        cc = mmul.mat_mul(aa, bb);
+        seconds = tt.time();
+
+        cout << "Second run took " << seconds << " seconds." << endl;
+    }
+
+
     if(!check) {
         cout << "Result not checked." << endl;
         return;
@@ -81,8 +99,9 @@ main(int argc, char* argv[])
 
     int random_size = 0;
     bool check_result = false;
+    bool print_timing = false;
 
-    while ((opt = getopt(argc, argv, "hpa:b:o:n:c")) != -1) {
+    while ((opt = getopt(argc, argv, "hpta:b:o:n:c")) != -1) {
         switch(opt) {
         case 'a':
             a_file = string(optarg);
@@ -102,6 +121,9 @@ main(int argc, char* argv[])
         case 'p':
             Bacon::use_opencl_cpu = true;
             break;
+        case 't':
+            print_timing = true;
+            break;
         case 'h':
             show_usage();
             return 0;
@@ -113,7 +135,7 @@ main(int argc, char* argv[])
 
 
     if (random_size != 0) {
-        random_test(random_size, check_result);
+        random_test(random_size, check_result, print_timing);
     }
     else {
         run_test(c_file, a_file, b_file);

### 2011-04-30 doc/Paper/Paper.tex
More progress on paper.

@@
 \documentclass{llncs}
+
 \usepackage{graphicx}
 \usepackage{float}
+\usepackage{verbatim}
+\usepackage{placeins}
+\usepackage{afterpage}
+
+\floatstyle{boxed}
+\newfloat{listing}{thp}{lop}
+\floatname{listing}{Listing}
 
@@
 
-This paper describes Bacon, a programming language that provides
-simpler development and better optimizations for OpenCL compatible
-general purpose GPUs.
+This paper describes Bacon, a programming language that makes it
+easier to write high performance kernels to run on for OpenCL
+compatible general purpose GPUs. The design and implementation is
+described, and benchmarks are given showing performance gains over
+comparable OpenCL programs.
 
@@
 
-Bacon was developed to try to improve the OpenCL programming
-experience. It extends the OpenCL C programming language and provides
-a compiler that hopes to solve these issues. C++ wrapper code is
-generated automatically. Just in time specialization allows the 
-illusion of dynamic stack allocation as well as the unrolling of
-some loops.
+Bacon was developed to try to improve upon the OpenCL programming
+experience and make it easier to write high performance programs for
+GPU hardware. It does this by extending the OpenCL C syntax into a new
+language called Bacon C, and by performing some pre-optimizations as
+the Bacon C code is compiled to OpenCL C.
 
-Partial evaluation has been shown to provide improved performance in
-scientific computing applications. A good example of this is shown by Berlin
-\cite{Berlin:1990}.
+The key optimization performed by the Bacon compiler is just in time
+specialization. When a kernel is called, a specialzied version of the
+kernel is generated with some of its variable arguments replaced with
+compile-time constants. This allows for a variety of agressive
+optimizatons to be performed by the OpenCL compiler. Further, it
+allows for some variable sized array declarations to be turned into
+constant sized arrays that are allowed by OpenCL.
 
-The design of the OpenCL runtime system depends on device-specific
-code generation occuring at runtime, which suggests the use of
-parameter information for optimization at runtime.
+We evaluate these improvements below using matrix multiplication as a
+test case. Both simpler code and improved performance are
+demonstrated.
 
@@
 
-The Bacon programming language is based on OpenCL C with a small
-number of modifications:
+\begin{listing}[tb]
+\begin{verbatim}
+kernel
+Array2D<float>
+mat_mul(Array2D<float> aa, Array2D<float> bb) 
+{
+ SETUP:
+    global Array2D<float> cc[aa.rows, bb.cols];
 
-- Kernel definition format that allows auto-generation of
-  C++ interface code.
-- Native 2D and 3D array syntax.
-- Runtime value specialization, allowing for
-  - Variable size array declarations
-  - The expectation of loop unrolling
+ BODY:
+    @range [cc.rows, cc.cols];
 
-Because private array items are actually assigned to device registers,
-at least in the AMD implementation of OpenCL, this means that a loop
-over a variable sized local array can be expected to be as efficient
-as the manually unrolled version.
+    float sum = 0.0;
+    assert(aa.cols == bb.rows, 
+        "Matrixes must have compatible dimensions.");
+    for (int kk = 0; kk < aa.cols; ++kk) {
+        sum += aa[$row, kk] * bb[kk, $col];
+    }
+    cc[$y, $x] = sum;
+    return cc;
+}
+\end{verbatim}
+\caption{Naive Matrix Multiplication in Bacon C}\label{mmk}
+\end{listing}
+
+Bacon C is based on OpenCL C with extensions for improved usability
+and to enable the automatic generation of C++ wrapper code. A sample
+Bacon C kernel that performs matrix multiplication is shown in
+Listing~\ref{mmk}.
+
+Wrapper code generation is enabled by the separation of each kernel
+declarations into separate SETUP and BODY sections. Conceptually, the
+SETUP section is for code is independent of any specific parallel
+thread while the BODY section is the code that runs in parallel. This
+allows the declaration and dynamic sizing of variables that will
+reside in GPU global memory and can be returned to the host process.
+
+Each BODY includes a @range declaration that specifes the range it
+will be executed over in parallel. Within the BODY, the current
+position in that range is held in special variables named {\tt\$row},
+{\tt\$col}, and {\tt\$dep} for the first, second, and third dimension
+respectively.
+
+Parameterized types for 1D, 2D and 3D arrays are provided
+natively. The types are parameterized using C++-style angle bracket
+syntax. Both declarations and element access use a comma separated
+list of dimensions in square brackets. The dimensions of these arrays
+can be accessed using struct-style dot notation.
+
+Additional error handling is provided through the {\tt assert} and
+{\tt fail} keywords which will raise exceptions in the host process if
+triggered. Fail throws an error if it is executed at all, while assert
+is triggered if its condition is false.
+
+Each Bacon kernel has a set of specialization variables. The
+dimensions of any arrays passed as arguments to a kernel are always
+specialization variables. Other specialization variables can be
+specified explicitly by declaring arguments using the {\tt const}
+qualifier. Whenever a kernel is called with a new set of
+specialization variables a specialized version of that kernel is
+generated and executed.
+
+\begin{listing}[t!]
+\begin{verbatim}
+kernel
+Array2D<float>
+blocked_mat_mul_private(Array2D<float> aa, Array2D<float> bb, 
+                        const uint blksz)
+{
+ SETUP:
+    global Array2D<float> cc[aa.rows, bb.cols];
+
+ BODY:
+    @range [cc.rows / blksz, cc.cols / blksz];
+
+    private Array2D<float> sum[blksz, blksz];
+    int ii, jj, kk, gg;
+
+    for (ii = 0; ii < blksz; ++ii) {
+        for (jj = 0; jj < blksz; ++jj) {
+            sum[ii, jj] = 0.0;
+        }
+    }
+
+    int base_ii = $row * blksz;
+    int base_jj = $col * blksz;
+    int base_kk;
+
+    for (gg = 0; gg < aa.cols / blksz; ++gg) {
+        base_kk = gg * blksz;
+
+        for (ii = 0; ii < blksz; ++ii) {
+            for (jj = 0; jj < blksz; ++jj) {
+                for (kk = 0; kk < blksz; ++kk) {
+                    sum[ii, jj] += aa[base_ii + ii, base_kk + kk] * 
+                        bb[base_kk + kk, base_jj + jj];
+                }
+            }
+        }
+    }
+
+    for (ii = 0; ii < blksz; ++ii) {
+        for (jj = 0; jj < blksz; ++jj) {
+            cc[base_ii + ii, base_jj + jj] = sum[ii, jj];
+        }
+    }
+
+    return cc;
+}
+\end{verbatim}
+\caption{Blocked Matrix Multiplication in Bacon C}\label{bmmk}
+\end{listing}
+\afterpage{\FloatBarrier}
+
+This specialization, in addition to providing performance benefits,
+allows for the simulation of variable sized arrays in thread-local
+memory as long as the array size depends only on {\tt const} variables
+and array dimensions. The blocked matrix multiply kernel in
+Listing~\ref{bmmk} gives an example of this feature.
 
@@
 
-Bacon consists of a compiler that translates Bacon C into an
-intermediate representation and generates interface code to load and
-execute that code at runtime from a C++ host application.
+The Bacon system consists of two pieces: the front-end compiler and the
+runtime library. The front-end library parses the Bacon C source and
+outputs a C++ wrapper and an intermediate form of the parsed code. The
+runtime library is called from the generated wrapper to load the
+intermediate form code, generate specialized OpenCL C code when a
+kernel is called, and run that code on the GPU using the OpenCL
+runtime.
+
+The system is built using Perl and C++. The front-end compiler parses
+the source code using Parse::Yapp, a yacc-compatible parser generator
+for Perl. This constructs abstract syntax tree, which is then walked
+to generate the C++ wrapper. The abstract syntax tree (AST) is then
+serialized for later loading by the runtime library.
+
+The generated C++ wrapper provides a C++ function with the kernel's
+type signature that can be called from the user's application. When
+this functions is called, the runtime library loads the AST and walks
+it to generate the specialized OpenCL code. Optimizatons are performed
+at code generation time without any intermediate pass that transforms
+the representation.
+
+The two optimizations that are performed by the Bacon library are
+constant propagation and loop unrolling. Constant propagation is used
+to determine the exact set and values of the specialized variables,
+building a mapping {\tt const} variables to their values so that
+references to these variables can be replaced with integer literals in
+the generated OpenCL C code. Loop unrolling is then performed on any
+loops that can be statically analized.
+
+Loop unrolling is included because neither the AMD nor Nvidia OpenCL
+compilers do it by default, and manually unrolling loops is time
+consuming and error prone. Loop unrolling in Bacon is an even more
+powerful optimization than usual due to the ability to either fully
+unroll loops or to unroll them by a factor that is guaranteed to
+evenly divide the iteration count because array sizes are always known
+due to just in time specialization.
+
+The implemetnation of Bacon is available publically under an open
+source license. The current version can be downloaded from the public
+git repository\footnote{http://code.ferrus.net/compilers/bacon.git}.
 
@@
 
-To analyze the performance of the Bacon system, we performed a number
-of tests multiplying random 4096x4096 matrixes of single precision
-point numbers.
-
-Testing was performed using the hardware shown in Table~\ref{hw}. At
-the time of this writing this represents an inexpensive (\~\$500) new
-PC for computer gaming, but both the CPU and GPU archetectures are
-less than one design generation behind the latest hardware from AMD.
-
-\begin{table}[htb]
-\begin{tabular}{ l @{\hspace{10pt}} l @{\hspace{10pt}} l }
-Device & Speed & RAM \\
-\hline
-\noalign{\smallskip}
-AMD Phenom II X3 720 & 3 Cores @ 2.8 GHz & 4GB DDR2 @ 800MHz \\
-\noalign{\smallskip}
-ATI Radeon 5830 & 1120 Stream Cores @ 800 MHz & 1GB GDDR5 @ 1000 MHz \\
-\noalign{\smallskip}
-\end{tabular}
-\caption{Hardware used for performance testing}\label{hw}
-\end{table}
-
-The baseline measurement is a naive matrix multiplication written in OpenCL C.
-
-\begin{table}
+\begin{table}[t!]
 \begin{tabular}{ l @{\hspace{10pt}} r @{\hspace{10pt}} r }
@@
 
-\begin{figure}[htb]
+In order to evaluate the performance of the Bacon system, we compare
+the run time of matrix multiplication kernels in both Bacon C and
+OpenCL C.
+
+Testing was performed on an ATI Radeon 5830 GPU. This is a cut down
+version of AMD's current top of the line GPU chipset, which at the
+time of this writing sells new for around \$140. A similar Nvidia card
+was purchased for comparison testing, but the comparsion was not
+performed due to implementation incompatibilities that we have not yet
+resolved.
+
+Four implementations of matrix multiplication were tested:
+
+\begin{description}
+  \item[Bacon - Naive] Shown in Listing~\ref{mmk}.
+  \item[OpenCL - Naive] An equivilent OpenCL C implementation.
+  \item[Bacon - Blocked] Shown in listing~\ref{bmmk}.
+  \item[OpenCL - Hand Vectorized] Hand unrolled using vector types to
+    compute 4x4 blocks. Based on a sample from the AMD OpenCL SDK.
+\end{description}
+
+Tests were performed on 4096x4096 matrices to create large enough
+chunks of work to be able to time the results without worrying about
+factors like OpenCL queue latencies. As table~\ref{mm1} shows, the
+optimizations provided by the Bacon system provide significant
+speedups on this test. As a comparison to show how good GPUs are for
+this kind of taask, a naive serial implementation running on a CPU can
+take as much as 45 minutes to do this calculation\footnote{Tested on
+  an AMD Phenom II X3 @ 2.8 GHz}, although a highly optimized
+implementation can approach the GPU times\footnote{ATLAS BLAS clocks
+  in at 7.8 seconds}.
+
+\begin{figure}[tbh]
 \begin{center}
@@
 \end{center}
-\caption{Summary of 4k Matrix Multiplication Performance}\label{unroll}
+\caption{Performance of 4k Matrix Multiplication with Unrolled Loops}\label{unroll}
 \end{figure}
@@
 \section{Conclusion}
-
-Hai
+Hai \cite{Berlin:1990}.
 

### 2011-05-03 doc/Paper/Paper.tex
Tests pass again.

@@
 described, and benchmarks are given showing performance gains over
-comparable OpenCL programs.
+comparable programs written directly in OpenCL C.
 
@@
 
-The OpenCL standard consists of two major pieces. First, it defines a
-programming language called OpenCL C for writing compute kernels to
-run on parallel hardware. Second, it defines runtime APIs for C and
-C++ that allow these kernels to be compiled, loaded, and executed from
-programs running on a host machine.
+The OpenCL standard\cite{opencl} consists of two major pieces. First,
+it defines a programming language called OpenCL C for writing compute
+kernels to run on parallel hardware. Second, it defines runtime APIs
+for C and C++ that allow these kernels to be compiled, loaded, and
+executed from programs running on a host machine.
 
@@
 The key optimization performed by the Bacon compiler is just in time
-specialization. When a kernel is called, a specialzied version of the
-kernel is generated with some of its variable arguments replaced with
-compile-time constants. This allows for a variety of agressive
-optimizatons to be performed by the OpenCL compiler. Further, it
-allows for some variable sized array declarations to be turned into
-constant sized arrays that are allowed by OpenCL.
+specialization. Rather than performing When a kernel is called, a
+specialzied version of the kernel is generated with some of its
+variable arguments replaced with compile-time constants. This allows
+for a variety of agressive optimizatons to be performed by the OpenCL
+compiler. Further, it allows for some variable sized array
+declarations to be turned into constant sized arrays that are allowed
+by OpenCL.
 
@@
 demonstrated.
+
+\section{Previous Work}
+
+Specialization, also known as partial evaluation, has been shown to be
+a very effective optimization by projects such as
+Tempo\cite{consel:1998}, which does general partial evaluation of C
+code at compile time. Just in time specialization was popularized by
+its use in typing dynamic languages, as shown in the Self
+language\cite{chambers:1992}. Just in time specialization on values has
+been used in Prolog systems by Bolz\cite{bolz:2010}.
 
@@
 The system is built using Perl and C++. The front-end compiler parses
-the source code using Parse::Yapp, a yacc-compatible parser generator
-for Perl. This constructs abstract syntax tree, which is then walked
-to generate the C++ wrapper. The abstract syntax tree (AST) is then
-serialized for later loading by the runtime library.
+the source code using Parse::Yapp\cite{parse::yapp}, a yacc-compatible
+parser generator for Perl. This constructs abstract syntax tree, which
+is then walked to generate the C++ wrapper. The abstract syntax tree
+(AST) is then serialized for later loading by the runtime library.
 
@@
 the run time of matrix multiplication kernels in both Bacon C and
-OpenCL C.
+OpenCL C. These results are summarized in Table~\ref{mm1}, which shows
+that Bacon is able to provide measurable performance improvements over
+similar programs written directly in OpenCL C.
 
-Testing was performed on an ATI Radeon 5830 GPU. This is a cut down
-version of AMD's current top of the line GPU chipset, which at the
-time of this writing sells new for around \$140. A similar Nvidia card
-was purchased for comparison testing, but the comparsion was not
-performed due to implementation incompatibilities that we have not yet
-resolved.
+Testing was performed on an ATI Radeon HD 5830 GPU. This is a cut down
+version of AMD's current top of the line GPU, the Radeon HD 6970. The
+5830 sells for around \$140 at the time of this writing compared to
+\$340 for the 6970, while providing over 60\% of the theoretical
+compute power. A similar Nvidia card was purchased for comparison
+testing, but the comparsion was not performed due to implementation
+incompatibilities that we have not yet resolved.
 
@@
 
-Tests were performed on 4096x4096 matrices to create large enough
-chunks of work to be able to time the results without worrying about
-factors like OpenCL queue latencies. As table~\ref{mm1} shows, the
-optimizations provided by the Bacon system provide significant
-speedups on this test. As a comparison to show how good GPUs are for
-this kind of taask, a naive serial implementation running on a CPU can
-take as much as 45 minutes to do this calculation\footnote{Tested on
-  an AMD Phenom II X3 @ 2.8 GHz}, although a highly optimized
-implementation can approach the GPU times\footnote{ATLAS BLAS clocks
-  in at 7.8 seconds}.
+Timings were taken multiplying 4096x4096 matrices. Each test was
+performed five times and the average result was taken. The times were
+very consistent; most tests had a coefficient of variation under one
+percent. The speedups over the naive OpenCL implementationa are shown
+in Figure~\ref{unroll}.
 
@@
 
+These measurements clearly show that value specialization provides
+significant speedups over a non-specialized OpenCL kernel. This result
+was expected; providing constant values at compile time allows the
+OpenCL C compiler to do extensive and agressive constant
+propagantion-based optimizations.
+
+The results from loop unrolling are more complicated. Unrolling loops
+too much drastically decrease performance, most likely due to the
+exhaustion of registers on the GPU device. Still, when properly tuned,
+loop unrolling provides significant speedups allowing the blocked
+Bacon kernel to beat the hand vectorized OpenCL kernel by nearly 30
+percent.
+
+\section{Future Work}
+
+The Bacon system could benefit from a number of extensions to improve
+its performance, generality, and utility to developers. One obvious
+area of improvement would be to automatically determine the best
+factor for loop unrolling rather than expecting it to be tuned
+manually. The unrolling analysis could also be extended to the
+implicit parallel ``loops'' created by the range of the computation;
+optimally, it should be possible to generate kernels like the blocked
+matrix mulitplication shown from the naive matrix multiplication
+kernel. Another simple improvement would be to perform strip mining
+(Wolfe \cite{wolfe:1996}, section 9.8) and generate explicit vector
+code for loops.
+
+Just in time specialization as a method to improve performance on
+parallel computations seems very promising in general, but there are
+some limitations imposed by the use of OpenCL as a compiler target. An
+implementation of this technique targetting a parallel processor at a
+lower level would allow for more flexibiltiy and much faster
+specialization times.
+
 \section{Conclusion}
-Hai \cite{Berlin:1990}.
+
+We have shown that Bacon allows a high performance GPU matrix
+multiplication kernel to be written in a naive style and executed with
+nearly the performance of a hand-vectorized OpenCL kernel due to the
+performance benefits of just in time specialization. Further, we have
+shown that by re-writing that kernel with a generalized block strategy
+and selecting appropriate loop unrolling settings, the performance of
+the hand vectorized OpenCL kernel can be exceeded.
+
+We are distributing this tool in the hope that it will be practically
+useful for developing kernels targeting OpenCL compatible GPU devices.
 

### 2011-05-29 doc/Paper/Paper.tex
Paper abstract.

@@
 
+% Keywords:
+%   data parallel, gpu, opencl, jit specialization
+
 \begin{abstract}
 
-This paper describes Bacon, a programming language that makes it
-easier to write high performance kernels to run on for OpenCL
-compatible general purpose GPUs. The design and implementation is
-described, and benchmarks are given showing performance gains over
-comparable programs written directly in OpenCL C.
+This paper describes Bacon, a data-parallel programming system
+targetting OpenCL-compatible graphics processors. This system is built
+upon the existing OpenCL standard in order to make it easier for
+programmers to write high performance kernels for GPU accelerated
+applications. The OpenCL C syntax is extended into a new language,
+Bacon C, intended to make development significantly more convienient
+and enabling pre-optimizations based on just-in-time specialization as
+this code is compiled via OpenCL at runtime.
+
+Benchmarks are provided for matrix multiplication comparing a Bacon
+implementation to versions using OpenCL directly. Speedups are
+demonstrated both for naive implementations and when comparing a naive
+Bacon implementation to a hand-vectorized OpenCL kernel. This latter
+result demonstrates the benefit of the total loop unrolling enabled by
+just-in-time specialization. Additionally, a Bacon implementation of a
+more complex computation, stereo disparity, is considered.
 

### 2011-05-30 doc/Paper/Paper.tex
Paper tweaks.

@@
 This paper describes Bacon, a data-parallel programming system
-targetting OpenCL-compatible graphics processors. This system is built
+targeting OpenCL-compatible graphics processors. This system is built
 upon the existing OpenCL standard in order to make it easier for
@@
 applications. The OpenCL C syntax is extended into a new language,
-Bacon C, intended to make development significantly more convienient
+Bacon C, intended to make development significantly more convenient
 and enabling pre-optimizations based on just-in-time specialization as
@@
 years. In response to the platform specific programming solutions from
-Nvidia and Microsoft, Apple developed the OpenCL standard as an open and
-cross platform programming interface to this hardware.
+Nvidia and Microsoft, Apple developed the OpenCL standard as an open
+and cross platform programming interface to this hardware. This
+standard has since been implemented by a number of major vendors
+including AMD, Nvidia, Intel, and IBM.
 
@@
 OpenCL C uses the syntax of C99 and provides a set of built in data
-types and functions that expose the numeric computatation capabilities
+types and functions that expose the numeric computation capabilities
 common to modern GPU devices. The specification explicitly disallows
@@
 Rather than providing a stand-alone program to compile OpenCL C
-kernels, the OpenCL C and C++ APIs give developers the nessisary
-pieces to build a compiler into their host program. This allows OpenCL
-programs to be portable across different hardware archetectures by
+kernels, the OpenCL C and C++ APIs give developers the pieces
+necessary to build a compiler into their host program. This allows OpenCL
+programs to be portable across different hardware by architectures
 delaying compilation until runtime when the target GPU device is
 known, but requires each developer to write quite a bit of code to
-load the source code and perform various other bookkeeping.
+load the source code and perform various other bookkeeping activities.
 
-Bacon was developed to try to improve upon the OpenCL programming
+Bacon was developed to improve upon the OpenCL programming
 experience and make it easier to write high performance programs for
 GPU hardware. It does this by extending the OpenCL C syntax into a new
-language called Bacon C, and by performing some pre-optimizations as
+language called Bacon C, and by performing pre-optimizations as
 the Bacon C code is compiled to OpenCL C.
@@
 The key optimization performed by the Bacon compiler is just in time
-specialization. Rather than performing When a kernel is called, a
-specialzied version of the kernel is generated with some of its
-variable arguments replaced with compile-time constants. This allows
-for a variety of agressive optimizatons to be performed by the OpenCL
-compiler. Further, it allows for some variable sized array
-declarations to be turned into constant sized arrays that are allowed
-by OpenCL.
+specialization. When a Bacon kernel is written, some integer arguments
+can be marked as specialization-time constants. When a kernel is first
+called with a given set of values for those arguments a specialized
+version of that kernel is generated with those variable arguments
+replaced with constant values. This allows for a variety of
+optimizations be performed by the OpenCL compiler. Further, it allows
+for any variable sized array declarations that depend only on these
+specialization-time constants to be turned into constant sized arrays
+that are allowed by OpenCL.
 
-We evaluate these improvements below using matrix multiplication as a
-test case. Both simpler code and improved performance are
-demonstrated.
+These improvements are evaluated using matrix multiplication as a test
+case. Both simpler code and improved performance are demonstrated.
 
-\section{Previous Work}
+\section{Previous and Similar Work}
+
+\subsection{Partial Evaluation}
 
@@
 been used in Prolog systems by Bolz\cite{bolz:2010}.
+
+\subsection{OpenCL Libraries and Front-Ends}
+
+Bacon is not the first attempt to provide an improved programmer
+experience for OpenCL.
 
@@
     assert(aa.cols == bb.rows, 
-        "Matrixes must have compatible dimensions.");
+        "Matrices must have compatible dimensions.");
     for (int kk = 0; kk < aa.cols; ++kk) {
@@
 
-Each BODY includes a @range declaration that specifes the range it
+Each BODY includes a @range declaration that specifies the range it
 will be executed over in parallel. Within the BODY, the current
@@
 
-Parameterized types for 1D, 2D and 3D arrays are provided
-natively. The types are parameterized using C++-style angle bracket
+Parametrized types for 1D, 2D and 3D arrays are provided
+natively. The types are parametrized using C++-style angle bracket
 syntax. Both declarations and element access use a comma separated
@@
 this functions is called, the runtime library loads the AST and walks
-it to generate the specialized OpenCL code. Optimizatons are performed
+it to generate the specialized OpenCL code. Optimizations are performed
 at code generation time without any intermediate pass that transforms
@@
 
-The implemetnation of Bacon is available publically under an open
+The implementation of Bacon is available publicly under an open
 source license. The current version can be downloaded from the public
@@
 compute power. A similar Nvidia card was purchased for comparison
-testing, but the comparsion was not performed due to implementation
+testing, but the comparison was not performed due to implementation
 incompatibilities that we have not yet resolved.
@@
   \item[Bacon - Naive] Shown in Listing~\ref{mmk}.
-  \item[OpenCL - Naive] An equivilent OpenCL C implementation.
+  \item[OpenCL - Naive] An equivalent OpenCL C implementation.
   \item[Bacon - Blocked] Shown in listing~\ref{bmmk}.
@@
 very consistent; most tests had a coefficient of variation under one
-percent. The speedups over the naive OpenCL implementationa are shown
+percent. The speedups over the naive OpenCL implementation are shown
 in Figure~\ref{unroll}.
@@
 optimally, it should be possible to generate kernels like the blocked
-matrix mulitplication shown from the naive matrix multiplication
+matrix multiplication shown from the naive matrix multiplication
 kernel. Another simple improvement would be to perform strip mining
@@
 some limitations imposed by the use of OpenCL as a compiler target. An
-implementation of this technique targetting a parallel processor at a
-lower level would allow for more flexibiltiy and much faster
+implementation of this technique targeting a parallel processor at a
+lower level would allow for more flexibility and much faster
 specialization times.

### 2011-05-30 doc/Paper/Paper.tex
Bacon paper: Now with less fail.

@@
 implementation to versions using OpenCL directly. Speedups are
-demonstrated both for naive implementations and when comparing a naive
-Bacon implementation to a hand-vectorized OpenCL kernel. This latter
-result demonstrates the benefit of the total loop unrolling enabled by
-just-in-time specialization. Additionally, a Bacon implementation of a
-more complex computation, stereo disparity, is considered.
+demonstrated both for naive implementations and when comparing a Bacon
+implementation of generalized block decomposed matrix multiplication
+to a hand-vectorized OpenCL kernel. This latter result demonstrates
+the benefit of the total loop unrolling enabled by just-in-time
+specialization. Additionally, a Bacon implementation of a more complex
+computation, stereo disparity, is considered.
 
@@
 Bacon is not the first attempt to provide an improved programmer
-experience for OpenCL.
+experience for OpenCL. Rick Webber has developed an improved C++ API
+called clUtil\cite{webber:2011}. Bindings for other languages, like
+the JOCL\cite{hutter:2011} binding for Java, provide APIs at a variety
+of levels of abstraction.
+
+Another very interesting approach allows the programmer to describe
+the computation to be executed on the GPU in the same language as the
+rest of the program. An OpenCL kernel is then generated
+automatically. This has the potential benefit of increased expressive
+power and programmer familiarity at the cost of increased complexity
+and necessarily leaky abstraction when the host language contains
+features that can't be cleanly transformed into GPU code. Examples of
+this approach include CLyther\cite{rossross:2011} for Python and
+ScalaCL\cite{chafik:2011} for the functional language Scala.
 
@@
     }
-    cc[$y, $x] = sum;
+    cc[$row, $col] = sum;
     return cc;
@@
 
-Parametrized types for 1D, 2D and 3D arrays are provided
+Parametrized types for 1D, 2D, and 3D arrays are provided
 natively. The types are parametrized using C++-style angle bracket
@@
 memory as long as the array size depends only on {\tt const} variables
-and array dimensions. The blocked matrix multiply kernel in
+and array dimensions. This works around one of the main constraints
+imposed by OpenCL. The blocked matrix multiply kernel in
 Listing~\ref{bmmk} gives an example of this feature.
@@
 
-The Bacon system consists of two pieces: the front-end compiler and the
-runtime library. The front-end library parses the Bacon C source and
-outputs a C++ wrapper and an intermediate form of the parsed code. The
-runtime library is called from the generated wrapper to load the
-intermediate form code, generate specialized OpenCL C code when a
-kernel is called, and run that code on the GPU using the OpenCL
-runtime.
+The Bacon system consists of two pieces: the compiler and the runtime
+library. The compiler parses the Bacon C source and outputs a C++
+wrapper and the serialized abstract syntax tree (AST). The runtime
+library is called from the generated wrapper to load the intermediate
+form code, generate specialized OpenCL C code when a kernel is called,
+and run that code on the GPU using the OpenCL runtime.
 
@@
 the source code using Parse::Yapp\cite{parse::yapp}, a yacc-compatible
-parser generator for Perl. This constructs abstract syntax tree, which
-is then walked to generate the C++ wrapper. The abstract syntax tree
-(AST) is then serialized for later loading by the runtime library.
+parser generator for Perl. This constructs the abstract syntax tree,
+which is then walked to generate the C++ wrapper.
 
@@
 The two optimizations that are performed by the Bacon library are
-constant propagation and loop unrolling. Constant propagation is used
-to determine the exact set and values of the specialized variables,
-building a mapping {\tt const} variables to their values so that
-references to these variables can be replaced with integer literals in
-the generated OpenCL C code. Loop unrolling is then performed on any
-loops that can be statically analized.
+constant propagation and loop unrolling. Constant propagation
+calculates the values of all the variables that have been marked as
+\tt{const} by the programmer. If the value of any of these variables
+cannot be computed from the specialized arguments to the kernel, the
+Bacon runtime will throw an exception.  This information is used to
+construct a symbol table, and references to these variables are
+replaced with their constant integer values in the generated OpenCL
+code. Loop unrolling is then performed on any loops that can be
+statically analyzed.
 
@@
 In order to evaluate the performance of the Bacon system, we compare
-the run time of matrix multiplication kernels in both Bacon C and
-OpenCL C. These results are summarized in Table~\ref{mm1}, which shows
-that Bacon is able to provide measurable performance improvements over
-similar programs written directly in OpenCL C.
+the run time of matrix multiplication kernels written in both Bacon C
+and OpenCL C. These results are summarized in Table~\ref{mm1}, which
+shows that Bacon is able to provide measurable performance
+improvements over similar programs written directly in OpenCL C.
 
@@
   \item[OpenCL - Naive] An equivalent OpenCL C implementation.
-  \item[Bacon - Blocked] Shown in listing~\ref{bmmk}.
+  \item[Bacon - Blocked] Shown in Listing~\ref{bmmk}.
   \item[OpenCL - Hand Vectorized] Hand unrolled using vector types to
@@
 
-Timings were taken multiplying 4096x4096 matrices. Each test was
-performed five times and the average result was taken. The times were
-very consistent; most tests had a coefficient of variation under one
-percent. The speedups over the naive OpenCL implementation are shown
-in Figure~\ref{unroll}.
+The execution time of these four kernels was tested on randomly
+generated 4096x4096 matrices. Each test was performed five times and
+the average result was taken. The times were very consistent; most
+tests had a coefficient of variation under one percent. The speedups
+over the naive OpenCL implementation are shown in Figure~\ref{unroll}.
 
@@
 significant speedups over a non-specialized OpenCL kernel. This result
-was expected; providing constant values at compile time allows the
-OpenCL C compiler to do extensive and agressive constant
-propagantion-based optimizations.
+can be explained by the fact taht providing constant values at compile
+time allows the OpenCL C compiler to do extensive constant
+propagation-based optimizations.
 
@@
 percent.
+
+\section{Stereo Disparity in Bacon}
+
+In order to demonstrate that Bacon is suitable for complex
+applications we consider an implementation of stereo disparity
+matching. This is an application that is well suited to GPU
+implementation due to being computationally expensive and naturally
+parallel. The stereo disparity computation is significantly more
+complex than matrix multiplication, requiring multiple steps such that
+the entire process can not easily be implemented as a single kernel.
+
+The stereo disparity matching problem is to generate a disparity image
+given a pair of images from a stereo camera rig. Each pixel ($y,x$) in
+the disparity image has a value ($v$) specifying a pixel offset that
+matches the pixel at ($y,x$) in the left image to the pixel at
+($y,x+v$) in the right image. These matched pixels correspond to the
+same physical point in scene viewed by the cameras. This disparity
+image can then be used to compute depth information and create a 3D
+point cloud describing the scene, which is useful in a variety of
+practical applications like robotics.
+
+The stereo disparity method implemented used the Census
+transform\cite{zabih:1994} both with simple local window matching and
+with a partial implementation of the semi-global matching (SGM) technique
+described by Hirschmuller\cite{hirschmuller:2005}. Testing during
+development was done on sample image pairs from the Middlebury 2005
+Stereo Dataset\cite{hirschmuller:2005}. Both a CPU implementation and
+a Bacon implementation were written.
+
+Although a complete implementation of the SGM algorithm was not
+completed and neither Bacon implementation has been properly
+optimized, the results are promising. The simple local window matcher
+in Bacon produces the same output as the CPU version. The partial SGM
+implementation produces a more accurate disparity image but takes
+longer to run.
+
+Including specialization time, the Bacon version of the local window
+matcher takes almost as much time as the CPU version, but on
+subsequent calls it takes less than 0.2 seconds which is more than a
+10x speedup. This provides a good example of the practicality of the
+value specialization mechanism; for processing video an extra
+computation cost on the first frame will be quickly amortized over
+subsequent frames.
 

### 2011-08-16 examples/GoodStereo/stereo.cc
Bi-directional 1/2 SGM.

@@ -26,25 +26,20 @@ stereo_disparity(Stereo& ss, cv::Mat matL, cv::Mat matR)
     Bacon::Timer tt_full;
 
     // convert inputs to Array2D
-    // scale
-    // census 1/2
-    // SGM 1/2
-    // consistency 1/2
-    // fill unknown pixels magically
-    // census full
-    // restricted range matching
-    // consistency check
-
     Bacon::Timer tt_conv;
     Image2D<cl_uchar> aL = mat_to_image2d<cl_uchar>(matL, 8);
     Image2D<cl_uchar> aR = mat_to_image2d<cl_uchar>(matR, 8);
     cout << "Conversion: " << tt_conv.time() << endl;
 
+    // scale
     Bacon::Timer tt_scale;
     Image2D<cl_uchar> hL = ss.scale_half(aL);
     Image2D<cl_uchar> hR = ss.scale_half(aR);
     cout << "Scale: " << tt_conv.time() << endl;    
 
+    // census 1/2
+    // SGM 1/2
+    // consistency 1/2
     Bacon::Timer tt_ch;
     Image2D<cl_ulong> chL = ss.sparse_census(hL);
     Image2D<cl_ulong> chR = ss.sparse_census(hR);
@@ -52,16 +47,16 @@ stereo_disparity(Stereo& ss, cv::Mat matL, cv::Mat matR)
 
     Array3D<cl_uchar> pL(8, chL.rows(), chL.cols());
 
-    Bacon::Timer tt_sgm_h;
+    Bacon::Timer tt_sgm_hL;
     ss.sgm_h(pL, chL, chR, +1);
-    cout << "SGM H: " << tt_sgm_h.time() << endl;
+    cout << "SGM HL: " << tt_sgm_hL.time() << endl;
 
     //show_pspace_slice("pL[0]", pL, 0);
     //show_pspace_slice("pL[1]", pL, 1);
 
-    Bacon::Timer tt_sgm_v; 
+    Bacon::Timer tt_sgm_vL; 
     ss.sgm_v(pL, chL, chR, +1);
-    cout << "SGM V: " << tt_sgm_v.time() << endl;
+    cout << "SGM VL: " << tt_sgm_vL.time() << endl;
 
     //show_pspace_slice("pL[2]", pL, 2);
     //show_pspace_slice("pL[3]", pL, 3);
[…]

### 2011-08-18 examples/FastStereo/show_image.cc
Working full 8-way half-resolution SGM implementation.

@@ -1,4 +1,6 @@
 
+#include <boost/scoped_array.hpp>
+
 #include "show_image.hh"
 
 inline int
@@ -8,7 +10,7 @@ min(int aa, int bb)
 }
 
 void
-show_image(const char* title, cv::Mat image)
+show_image(const char* title, cv::Mat image, int scale)
 {
     cv::namedWindow(title, CV_WINDOW_AUTOSIZE);
     cv::imshow(title, image);
@@ -28,4 +30,52 @@ show_census(const char* title, uint64_t* data, int rows, int cols)
     }
 
     show_image(title, image);
+}
+
+void 
+show_census_a2d(const char* title, Bacon::Array2D<cl_ulong>& aa)
+{
+    show_census(title, aa.ptr(), aa.rows(), aa.cols());
+}
+
+void
+show_pspace_slice(const char* title, Bacon::Array3D<cl_uchar>& pspace, int slice, int scale)
+{
+    cv::Mat image(pspace.rows(), pspace.cols(), CV_8UC1);
+
+    for (int ii = 0; ii < pspace.rows(); ++ii) {
+        for (int jj = 0; jj < pspace.cols(); ++jj) {
+            image.at<uint8_t>(ii, jj) = scale * pspace.get(slice, ii, jj);
+        }
+    }
+
+    show_image(title, image);
+}
+
+void
+show_array2d(const char* title, Bacon::Array2D<cl_uchar>& aa, int scale)
+{
+    cv::Mat image(aa.rows(), aa.cols(), CV_8UC1);
+
+    for (int ii = 0; ii < aa.rows(); ++ii) {
+        for (int jj = 0; jj < aa.cols(); ++jj) {
+            image.at<uint8_t>(ii, jj) = aa.get(ii, jj);
+        }
+    }
+
+    show_image(title, image);    
+}
+
+void
+show_image2d(const char* title, Bacon::Image2D<cl_uchar>& aa, int scale)
+{
+    cv::Mat image(aa.rows(), aa.cols(), CV_8UC1);
+
+    for (int ii = 0; ii < aa.rows(); ++ii) {
+        for (int jj = 0; jj < aa.cols(); ++jj) {
+            image.at<uint8_t>(ii, jj) = scale * aa.get(ii, jj);
+        }
+    }
+
+    show_image(title, image);    
 }

### 2011-08-18 examples/GoodStereo/show_image.cc

@@ -10,7 +10,7 @@ min(int aa, int bb)
 }
 
 void
-show_image(const char* title, cv::Mat image)
+show_image(const char* title, cv::Mat image, int scale)
 {
     cv::namedWindow(title, CV_WINDOW_AUTOSIZE);
     cv::imshow(title, image);
@@ -39,13 +39,13 @@ show_census_a2d(const char* title, Bacon::Array2D<cl_ulong>& aa)
 }
 
 void
-show_pspace_slice(const char* title, Bacon::Array3D<cl_uchar>& pspace, int slice)
+show_pspace_slice(const char* title, Bacon::Array3D<cl_uchar>& pspace, int slice, int scale)
 {
     cv::Mat image(pspace.rows(), pspace.cols(), CV_8UC1);
 
     for (int ii = 0; ii < pspace.rows(); ++ii) {
         for (int jj = 0; jj < pspace.cols(); ++jj) {
-            image.at<uint8_t>(ii, jj) = pspace.get(slice, ii, jj);
+            image.at<uint8_t>(ii, jj) = scale * pspace.get(slice, ii, jj);
         }
     }
 
@@ -53,13 +53,27 @@ show_pspace_slice(const char* title, Bacon::Array3D<cl_uchar>& pspace, int slice
 }
 
 void
-show_array2d(const char* title, Bacon::Array2D<cl_uchar>& aa)
+show_array2d(const char* title, Bacon::Array2D<cl_uchar>& aa, int scale)
 {
     cv::Mat image(aa.rows(), aa.cols(), CV_8UC1);
 
     for (int ii = 0; ii < aa.rows(); ++ii) {
         for (int jj = 0; jj < aa.cols(); ++jj) {
             image.at<uint8_t>(ii, jj) = aa.get(ii, jj);
+        }
+    }
+
+    show_image(title, image);    
+}
+
+void
+show_image2d(const char* title, Bacon::Image2D<cl_uchar>& aa, int scale)
+{
+    cv::Mat image(aa.rows(), aa.cols(), CV_8UC1);
+
+    for (int ii = 0; ii < aa.rows(); ++ii) {
+        for (int jj = 0; jj < aa.cols(); ++jj) {
+            image.at<uint8_t>(ii, jj) = scale * aa.get(ii, jj);
         }
     }
 

### 2011-08-18 examples/GoodStereo/stereo.cc

@@ -22,6 +22,8 @@ cv::Mat
 stereo_disparity(Stereo& ss, cv::Mat matL, cv::Mat matR)
 {
     Bacon::Kernel::show_timing = true;
+    const int SGM_N   = 8;
+    const int SHOW_SC = 4;
 
     Bacon::Timer tt_full;
 
@@ -35,61 +37,84 @@ stereo_disparity(Stereo& ss, cv::Mat matL, cv::Mat matR)
     Bacon::Timer tt_scale;
     Image2D<cl_uchar> hL = ss.scale_half(aL);
     Image2D<cl_uchar> hR = ss.scale_half(aR);
-    cout << "Scale: " << tt_conv.time() << endl;    
+    cout << "Scale: " << tt_conv.time() << endl;
 
     // census 1/2
-    // SGM 1/2
-    // consistency 1/2
     Bacon::Timer tt_ch;
     Image2D<cl_ulong> chL = ss.sparse_census(hL);
     Image2D<cl_ulong> chR = ss.sparse_census(hR);
     cout << "Census Half: " << tt_ch.time() << endl;
 
+    // SGM 1/2 L
     Array3D<cl_uchar> pL(8, chL.rows(), chL.cols());
 
     Bacon::Timer tt_sgm_hL;
-    ss.sgm_h(pL, chL, chR, +1);
+    ss.sgm_h(pL, chL, chR, +1, SGM_N);
     cout << "SGM HL: " << tt_sgm_hL.time() << endl;
-
-    //show_pspace_slice("pL[0]", pL, 0);
-    //show_pspace_slice("pL[1]", pL, 1);
+    //show_pspace_slice("pL[0]", pL, 0, SHOW_SC);
+    //show_pspace_slice("pL[1]", pL, 1, SHOW_SC);
 
     Bacon::Timer tt_sgm_vL; 
-    ss.sgm_v(pL, chL, chR, +1);
+    ss.sgm_v(pL, chL, chR, +1, SGM_N);
     cout << "SGM VL: " << tt_sgm_vL.time() << endl;
+    //show_pspace_slice("pL[2]", pL, 2, SHOW_SC);
+    //show_pspace_slice("pL[3]", pL, 3, SHOW_SC);
+    
+    //show_pspace_slice("pl[6]", pL, 6, SHOW_SC);
+    //show_pspace_slice("pl[7]", pL, 7, SHOW_SC);
 
[…]

### 2011-08-18 examples/Stereo/FastStereo/show_image.cc
Rearrange stereo things.

@@ -1,0 +1,81 @@
+
+#include <boost/scoped_array.hpp>
+
+#include "show_image.hh"
+
+inline int
+min(int aa, int bb)
+{
+    return (aa < bb) ? aa : bb;
+}
+
+void
+show_image(const char* title, cv::Mat image, int scale)
+{
+    cv::namedWindow(title, CV_WINDOW_AUTOSIZE);
+    cv::imshow(title, image);
+    cv::waitKey(0);
+}
+
+void
+show_census(const char* title, uint64_t* data, int rows, int cols)
+{
+    cv::Mat image(rows, cols, CV_8UC1);
+    
+    for (int ii = 0; ii < rows; ++ii) {
+        for (int jj = 0; jj < cols; ++jj) {
+            uint64_t cv = data[ii*cols + jj];
+            image.at<unsigned char>(ii, jj) = min(255, 4 * cv);
+        }
+    }
+
+    show_image(title, image);
+}
+
+void 
+show_census_a2d(const char* title, Bacon::Array2D<cl_ulong>& aa)
+{
+    show_census(title, aa.ptr(), aa.rows(), aa.cols());
+}
+
+void
+show_pspace_slice(const char* title, Bacon::Array3D<cl_uchar>& pspace, int slice, int scale)
+{
+    cv::Mat image(pspace.rows(), pspace.cols(), CV_8UC1);
+
+    for (int ii = 0; ii < pspace.rows(); ++ii) {
+        for (int jj = 0; jj < pspace.cols(); ++jj) {
+            image.at<uint8_t>(ii, jj) = scale * pspace.get(slice, ii, jj);
+        }
+    }
+
+    show_image(title, image);
+}
+
+void
+show_array2d(const char* title, Bacon::Array2D<cl_uchar>& aa, int scale)
+{
+    cv::Mat image(aa.rows(), aa.cols(), CV_8UC1);
+
+    for (int ii = 0; ii < aa.rows(); ++ii) {
+        for (int jj = 0; jj < aa.cols(); ++jj) {
+            image.at<uint8_t>(ii, jj) = aa.get(ii, jj);
+        }
+    }
+
+    show_image(title, image);    
+}
+
+void
+show_image2d(const char* title, Bacon::Image2D<cl_uchar>& aa, int scale)
+{
+    cv::Mat image(aa.rows(), aa.cols(), CV_8UC1);
+
+    for (int ii = 0; ii < aa.rows(); ++ii) {
[…]

### 2011-08-18 examples/Stereo/FastStereo/stereo.cc

@@ -18,81 +18,66 @@ using std::string;
 
 using namespace Bacon;
 
-void
-show_pspace_slice(const char* title, Array3D<cl_uchar>& pspace, int slice)
-{
-    cv::Mat image(pspace.rows(), pspace.cols(), CV_8UC1);
-
-    for (int ii = 0; ii < pspace.rows(); ++ii) {
-        for (int jj = 0; jj < pspace.cols(); ++jj) {
-            image.at<uint8_t>(ii, jj) = pspace.get(slice, ii, jj);
-        }
-    }
-
-    show_image(title, image);
-}
-
-void
-show_array2d(const char* title, Array2D<cl_uchar>& aa)
-{
-    cv::Mat image(aa.rows(), aa.cols(), CV_8UC1);
-
-    for (int ii = 0; ii < aa.rows(); ++ii) {
-        for (int jj = 0; jj < aa.cols(); ++jj) {
-            image.at<uint8_t>(ii, jj) = aa.get(ii, jj);
-        }
-    }
-
-    show_image(title, image);    
-}
-
 cv::Mat
-stereo_disparity(cv::Mat matL, cv::Mat matR)
+stereo_disparity(Stereo& ss, cv::Mat matL, cv::Mat matR)
 {
-    Stereo ss;
-    //ss.ctx.show_timing = true;
+    Bacon::Kernel::show_timing = false;
+    Bacon::Timer tt_full;
 
-    Array2D<cl_uchar> aL = mat_to_array2d<cl_uchar>(matL, 8);
-    Array2D<cl_uchar> aR = mat_to_array2d<cl_uchar>(matR, 8);
+    // convert inputs to Array2D
+    //Bacon::Timer tt_conv;
+    Image2D<cl_uchar> aL = mat_to_image2d<cl_uchar>(matL, 8);
+    Image2D<cl_uchar> aR = mat_to_image2d<cl_uchar>(matR, 8);
+    //cout << "Conversion: " << tt_conv.time() << endl;
 
-    Array2D<cl_ulong> cL = ss.sparse_census(aL);
-    Array2D<cl_ulong> cR = ss.sparse_census(aR);
+    // scale
+    //Bacon::Timer tt_scale;
+    Image2D<cl_uchar> hL = ss.scale_half(aL);
+    Image2D<cl_uchar> hR = ss.scale_half(aR);
+    //cout << "Scale: " << tt_conv.time() << endl;
 
-#if 0
-    show_census("Census Left", cL.ptr(), cL.rows(), cL.cols());
[…]

### 2011-08-18 examples/Stereo/show_image.cc

@@ -1,31 +1,0 @@
-
-#include "show_image.hh"
-
-inline int
-min(int aa, int bb)
-{
-    return (aa < bb) ? aa : bb;
-}
-
-void
-show_image(const char* title, cv::Mat image)
-{
-    cv::namedWindow(title, CV_WINDOW_AUTOSIZE);
-    cv::imshow(title, image);
-    cv::waitKey(0);
-}
-
-void
-show_census(const char* title, uint64_t* data, int rows, int cols)
-{
-    cv::Mat image(rows, cols, CV_8UC1);
-    
-    for (int ii = 0; ii < rows; ++ii) {
-        for (int jj = 0; jj < cols; ++jj) {
-            uint64_t cv = data[ii*cols + jj];
-            image.at<unsigned char>(ii, jj) = min(255, 4 * cv);
-        }
-    }
-
-    show_image(title, image);
-}

### 2011-08-18 doc/Paper/Paper.tex
Some improvements to the paper.

@@
 
-\section{Previous and Similar Work}
+\section{Previous Work}
 
@@
 calculates the values of all the variables that have been marked as
-\tt{const} by the programmer. If the value of any of these variables
+{\tt const} by the programmer. If the value of any of these variables
 cannot be computed from the specialized arguments to the kernel, the
@@
 Loop unrolling is included because neither the AMD nor Nvidia OpenCL
-compilers do it by default, and manually unrolling loops is time
+compilers do it by default and manually unrolling loops is time
 consuming and error prone. Loop unrolling in Bacon is an even more
